---
title: "IRT calibration"
author: "Geraldo B. Padilla F."
date: "05-02-2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, fig.width = 5, fig.height = 3)
```

```{r packages, message=FALSE}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(lattice)
library(stats4)
library(difR)
library(ltm)
library(MASS)
library(msm)
library(xxIRT)
library(mirt)
library(shiny)
library(birtr)
```

```{r data.set}
setwd('C:/Users/Lenovo/OneDrive - University of Iowa/PhD portfolio/2. Courses/Spring 2022/PSQF62620001 Item Response Theory (Deborah Harris)/PSQF62620001-Item-Response-Theory/0.6 Calibration Exercise/')
data <- read.fwf("irt_data for calibration.txt", widths = c(4, rep(1,101)))
#data <- data[,1:101]
colnames(data) <- c('candidate', paste0('math', seq(1,60)), paste('read', seq(1,40)), 'Group_ID')
```

## 3PL calibration on 1000 examinees (rbind 0001 - 1000 examinees), 30 items math
```{r 3PL.2000math}
data1 <- data[1:1000, paste0("math",seq(1,30))]
m1 <- itemParEst(data1, model = "3PL", engine = "ltm")
m1 = data.frame(m1)

m3 = mirt(data1, 1, itemtype = '3PL')
m3_coef = data.frame(coef(m3))
write.csv(me_coef, 'm3_coef.csv')
m3.1 = read_csv2('m3_coef.csv')
data.frame(m3.1)
m3.1$u = NULL
coef(m3)
  
#fit rasch model (i.e., 1PL)
IRTrasch <- rasch(IQbin, IRT.param = TRUE,constraint = cbind(length(IQbin) + 1, 1))

#fit 2PL model
IRT2pl <- ltm(data1~z1,IRT.param = TRUE)

#fit 3PL model
IRT3pl <- tpm(data1,type="latent.trait",IRT.param = TRUE)
m2 = data.frame(coef(IRT3pl, prob = TRUE))

names(m2)
m1.1 = dplyr::select(m1, a, b, c)
m2.1 = m2%>%
  dplyr::rename(c = Gussng, b = Dffclt, a = Dscrmn)%>%
  dplyr::select(a, b, c)

anova(IRT2pl, IRT3pl)







#return item parameters for each model
coef(IRTrasch, prob = TRUE)
coef(IRT2pl, prob = TRUE)
coef(IRT3pl, prob = TRUE)

#compare model fit
anova(IRTrasch,IRT2pl)
anova(IRT2pl,IRT3pl)

#return goodness of fit of the items with chi-square test
item.fit(IRT2pl)

#Plot ICC's for each item
plot(IRT3pl,type="ICC")

#Plot ICC's for one item
plot(IRT3pl,type="ICC", items = 1)

#Plot Test Information Function 
plot(IRT3pl,type="IIC",items=0) 

#ability estimates
est<-factor.scores(IRT3pl)

est$score.dat$z1[1:5]

#test undimensionality
#first, fit 2PL model with 2 factors
IRT2pl_2 <- ltm(IQbin~z1+z2)

#second, run liklihood ratio test comparing 1 to 2 factor models
anova(IRT2pl,IRT2pl_2)

# the engine for estimating the 1PL model, either "ltm" (default) or "lme4"
#write.csv(par_2000p_60m_3PL,paste0(path,"par_2000p_60m_3PL.csv"))
write.csv(m_2000p_60math_3PL, file = "m_2000p_60math_3PL.csv")
```

## 3PL calibration on 2000 examinees (rbind 2001 - 4000 examinees), 60 item math
```{r}
df_2000p2_60math_3PL <- data[2001:4000, paste0("math",seq(1,60))]
m_2000p2_60math_3PL <- itemParEst(df_2000p2_60math_3PL, model = "3PL", engine = "ltm")
# the engine for estimating the 1PL model, either "ltm" (default) or "lme4"
#write.csv(par_2000p_60m_3PL,paste0(path,"par_2000p_60m_3PL.csv"))
write.csv(m_2000p2_60math_3PL, file = "m_2000p2_60math_3PL.csv")
```

## 2PL calibration on 2000 examinees (rbind 0001 - 2000 examinees), 60 item math
```{r 2PL.2000math}
m_2000p_60math_2PL <- itemParEst(df_2000p2_60math_3PL, model = "2PL",engine = "ltm")
#write.csv(m_2000p_60math_2PL,paste0(path,"m_2000p_60math_2PL.csv"))
write.csv(m_2000p_60math_2PL, file = "m_2000p_60math_2PL.csv")
```

---

## 3PL calibration on 2000 examinees (rbind 0001 - 2000 examinees), 60 item math
```{r 3PL.2000math}
df_2000p_60math_3PL <- data[1:2000, paste0("math",seq(1,60))]
m_2000p_60math_3PL <- itemParEst(df_2000p_60math_3PL, model = "3PL", engine = "ltm")
# the engine for estimating the 1PL model, either "ltm" (default) or "lme4"
#write.csv(par_2000p_60m_3PL,paste0(path,"par_2000p_60m_3PL.csv"))
write.csv(m_2000p_60math_3PL, file = "m_2000p_60math_3PL.csv")
```

---

```{r}
itemplot(shiny = T)
```

```{r}
icc(a = 1.3, b = -1.3, c = .10)
icc(a = .6, b = -0.1, c = .17)
icc(a = .7, b = .9, c = .18)
iccplot(a = .5621, b = 2.6633)
iccplot(a = 1.0286, b = 1.1785)
```






## 3PL calibration on 1000 examinees (rbind 1001 - 2000),  60 item math
```{r 3PL.1000math}
tmp1000m <- data[1001:2000,paste0("math",seq(1,60))]
par_1000p_60m_3PL <-itemParEst(tmp1000m,model = "3PL",engine = "ltm")
write.csv(par_1000p_60m_3PL, paste0(path, 'par_1000p_60m_3PL.csv'))
```

## 2PL calibration on 2000 examinees (rbind 0001 - 2000 examinees), 60 item math
```{r 2PL.2000math}
par_2000p_60m_2pl <- itemParEst(tmp2000m, model = "2PL",engine = "ltm")
write.csv(par_2000p_60m_2pl,paste0(path,"par_2000p_60math_2pl.csv"))
```

## 2PL calibration on 1000 examinees (rbind 1001 - 2000),  60 item math
```{r 2PL.1000math}
par_1000p_60m_2pl <-itemParEst(tmp1000m,model = "2PL",engine = "ltm")
write.csv(par_1000p_60m_2pl, paste0(path, 'par_1000p_60m_2pl.csv'))
```

----

### comparing item parameter using different sample size
```{r}
plot(par_2000p_60m[,"a"],par_1000p_60m[,"a"],main = "a parameter comparison: 1000 vs 2000")
abline(0,1,col="red")
```

```{r}
plot(par_2000p_60m[,"b"],par_1000p_60m[,"b"],main = "b parameter comparison: 1000 vs 2000")
abline(0,1,col="red")
```

```{r}
plot(par_2000p_60m[,"c"],par_1000p_60m[,"c"],main = "c parameter comparison: 1000 vs 2000")
abline(0,1,col="red")
```

```{r}
plot(par_2000p_60m[,"se(a)"],par_1000p_60m[,"se(a)"],main = "se(a) parameter comparison: 1000 vs 2000")
abline(0,1,col="red")
```

```{r}
plot(par_2000p_60m[,"se(b)"],par_1000p_60m[,"se(b)"],main = "se(b) parameter comparison: 1000 vs 2000")
abline(0,1,col="red")
```

```{r}
plot(par_2000p_60m[,"se(c)"],par_1000p_60m[,"se(c)"],main = "se(c) parameter comparison: 1000 vs 2000")
abline(0,1,col="red")
```

## miskey cases

```{r}
# Change 0/1 on one random item on math (change 1 to 8, 0 to 1, 8 to 0 or however you want). 
# Calibrate it using same code/2000 examinees as above and see if it calibrates and what item stats look like (hopefully bad!) 
# I would like whatever output shows this item is “bad”
tmp <- data[1:2000,c("candidate_Id",paste0("math",seq(1,60)))]
tmp$candidate_Id <- stringr::str_pad(tmp$candidate_Id,4,pad = "0")
for(i in paste0("math",8)){
  a <- ifelse(tmp[,i]==1,0,1)
  tmp[,i]<- a
}
write.table(tmp,"irt_miskeyed.dat",col.names = F, row.names = F, sep = "",quote = F)

tmp <- tmp[,-1]

par_2000p_60m_bad <-itemParEst(tmp,model = "3PL",engine = "ltm")
par_2000p_60m_2pl_bad <-itemParEst(tmp,model = "2PL",engine = "ltm")

```

### 3PL-based parameter comparision
```{r}
plot(par_2000p_60m[,"a"],par_2000p_60m_bad[,"a"], main = "3pl-based a parameter")
abline(0,1,col="red")
```

```{r}
plot(par_2000p_60m[,"b"],par_2000p_60m_bad[,"b"], main = "3pl-based b parameter")
abline(0,1,col="red")
```

```{r}
plot(par_2000p_60m[,"c"],par_2000p_60m_bad[,"c"], main = "3pl-based c parameter")
abline(0,1,col="red")
```

### 2PL-based parameter comparision

```{r}
plot(par_2000p_60m_2pl[,"a"],par_2000p_60m_2pl_bad[,"a"], main = "2pl-based a parameter")
abline(0,1,col="red")
```

```{r}
plot(par_2000p_60m_2pl[,"b"],par_2000p_60m_2pl_bad[,"b"], main = "2pl-based b parameter")
abline(0,1,col="red")
```

### 3PL results

```{r}

# Score 20 examinees using the item parameters from #1 above and their 0/1s. 
# 19 should be from 3001 to 6000—any 19- and use their math 0/1s. 
# For the 20th, pick another simulee in that range BUT use the last 60 0/1s . 
# Because this examinee is scored with ‘wrong’ item parameters, fit should look BAD

tmp <- data[3001:3020,paste0("math",seq(1,60))]

theta <- model_3pl_eap_scoring(u = as.matrix(tmp), a = par_2000p_60m[,1],
                       b = par_2000p_60m[,2],
                       c = par_2000p_60m[,3],D=1)

theta_good <- data.frame(theta = theta$t, se = theta$sd)


tmp1 <- data[3001:3019,paste0("math",seq(1,60))]
tmp3 <- data[3020,42:101]
colnames(tmp1)<- paste0("I",seq(1,60))
colnames(tmp3)<- paste0("I",seq(1,60))

tmp_bad <- rbind(tmp1,tmp3)
colnames(tmp_bad)<- NULL
rownames(tmp_bad)<- NULL
theta <- model_3pl_eap_scoring(u = as.matrix(tmp_bad), a = par_2000p_60m[,1],
                               b = par_2000p_60m[,2],
                               c = par_2000p_60m[,3],D=1)

theta_bad <- data.frame(theta = theta$t, se = theta$sd)
```

```{r}
theta_bad$id <- as.character(seq(1,20))
theta_sum <- cbind(theta_good, theta_bad)
colnames(theta_sum) <- c("theta_good","se_good","theta_bad","se_bad","id")

ggplot(theta_sum, aes(x = theta_good, y = theta_bad,label = id, colour = factor(id))) +
  geom_point()+geom_label() +theme(legend.position = "none")+
  geom_abline(intercept = 0, slope = 1) +
  ggtitle("3PL: theta comparison")

ggplot(theta_sum, aes(x = se_good, y = se_bad,label = id, colour = factor(id))) +
  geom_point()+geom_label() +theme(legend.position = "none")+
  geom_abline(intercept = 0, slope = 1) +
  ggtitle("3PL: se comparison")


write.csv(theta_sum, paste0(path,"theta_comparison_3PL.csv"),row.names = F)
```

### 2PL results

```{r eval=FALSE}

# Score 20 examinees using the item parameters from #1 above and their 0/1s. 
# 19 should be from 3001 to 6000—any 19- and use their math 0/1s. 
# For the 20th, pick another simulee in that range BUT use the last 60 0/1s . 
# Because this examinee is scored with ‘wrong’ item parameters, fit should look BAD
tmp <- data[3001:3020,paste0("math",seq(1,60))]

theta <- model_3pl_eap_scoring(u = as.matrix(tmp), a = par_2000p_60m_2pl[,1],
                       b = par_2000p_60m_2pl[,2],
                       c = 0,D=1)

theta_good <- data.frame(theta = theta$t, se = theta$sd)


tmp1 <- data[3001:3019,paste0("math",seq(1,60))]
tmp3 <- data[3020,42:101]
colnames(tmp1)<- paste0("I",seq(1,60))
colnames(tmp3)<- paste0("I",seq(1,60))

tmp_bad <- rbind(tmp1,tmp3)
colnames(tmp_bad)<- NULL
rownames(tmp_bad)<- NULL
theta <- model_3pl_eap_scoring(u = as.matrix(tmp_bad), a = par_2000p_60m_2pl[,1],
                               b = par_2000p_60m_2pl[,2],
                               c = 0,D=1)

theta_bad <- data.frame(theta = theta$t, se = theta$sd)
```

```{r eval=FALSE}
theta_bad$id <- as.character(seq(1,20))
theta_sum <- cbind(theta_good, theta_bad)
colnames(theta_sum) <- c("theta_good","se_good","theta_bad","se_bad","id")

ggplot(theta_sum, aes(x = theta_good, y = theta_bad,label = id, colour = factor(id))) +
  geom_point()+geom_label() +theme(legend.position = "none")+
  geom_abline(intercept = 0, slope = 1) +
  ggtitle("2PL: theta comparison")

ggplot(theta_sum, aes(x = se_good, y = se_bad,label = id, colour = factor(id))) +
  geom_point()+geom_label() +theme(legend.position = "none")+
  geom_abline(intercept = 0, slope = 1) +
  ggtitle("2PL: se comparison")


write.csv(theta_sum, paste0(path,"theta_comparison_2PL.csv"),row.names = F)
```